---
title: "GLOBLDECL_WERNERSBACH"
author: "Rico Kronenberg, Lisa Marie Oehlschlägel"
date: "3 Mai 2018"
output: html_document
---

### GLOBLDECL_WERNERSBACH - Global Declaration of Wernersbach (catchment area)
The Wernersbach is the catchment area which is used in this case to analyze the water balance. It is situated near Dresden in the Tharandt Forest.

### Content

* [Files](#files)
* [Auxiliary to run the programm](#auxiliary-to-run-the-programm)
* [Constants](#constants)
* [Parameters and variables](#parameters-and-variables)

### Files
The Parameters are from the files:

* CTemperateDeciduousForest.txt
* SCscl.txt
* LMcu.txt
* IDefault.txt
* Xdefault.txt
* FUni50.txt

The symbols in the next table are characterizing the constants, parameters and variables more specific. They are written in front of the explenation of them in their topics. 

Symbol|Description
------|----------------
*     |indicates an input from the parameter file
**    |indicates a variable in the data file or precip input file
***   |indicates a constant
****  |indicates a variable initialized or parameter set in the data file
&     |indicates a variable initialized in the parameter file
@     |indicates a calculated parameter (constant for run)
(none)|all others are variables
(none)|at end of variable name indicates integer variable
&     |at end of variable name indicates long integer variable
$     |at end of variable name indicates string variable
xxxxP |flow for precipitation interval (mm)
xxxxD |flow for day (mm)
xxxxM |flow for month (mm)
xxxxY |flow for year (mm)


### Auxiliary to run the programm
The following table includes some numbers and limits for numbers in the model:

Name          |Value|Description
--------------|-----|---------------------------------------------------------------
ML            |25   |number of values in layer parameter arrays, set to 25 for B90V4
gctMaxoutvars |60   |maximum number of output variables, an array dimension
gctGenoutv    |19   |number of variables in GENOUT form
gctIntoutv    |22   |number of variables in INTRNOUT form
gctLayoutv    |12   |number of variables in LAYOUT form
gvals         |8    |max number of graph output values
maxgraphs     |100  |-

Auxiliary explenations:

* A
    + parsed values from INSTRNG
* aparsed
    + parsed values from gsbparsestring, first subscript is 0
* bad
    + error indicator passed back to calling routine
* B90path$
    + path for running program, used for .INI and .TMP
* cancelsave As Boolean
    + true if save or print should be cancelled
* chkout(1 To gctMaxoutvars, 1 To 5) As Integer
    + 1 for output, 5 intervals; indicates which variables/interval selected; stored in .INI; commadec As Boolean
* dfiletype
    + data file type, 1-csv,2-decimal period, 3-decimal comma
* EV(1 To 25) As New frmevalout
    + EVAL windows
* FG(1 To maxgraphs) As New frmgraph
    + multiple instances of graph
* gnumber
    + graph number
* graphon
    + true if graph has been initialized
* inoutdir$
    + latest input-output directory
* intrvl
    + output interval index, 1 ANN, 2 MON, 3 DAY, 4 PRE, 5 ITR
* ivar(1 To gctMaxoutvars), pvar(1 To gctMaxoutvars), dvar(1 To gctMaxoutvars), mvar(1 To gctMaxoutvars), yvar(1 To gctMaxoutvars)
    + list of var numbers for output, does not count mulltiple soil layers
* msg As String
* noenderror As Integer
    + 0 if no END in parameter file, -1 if END is found
* noruncontinue
    + 1 to prevent Run-Continue after soil parameter changes
* NLN As String
    + new line
* outfilestatus(1 To 5) As Integer
    + 0 if no output, 1 if output, subscript 1 ann, 2 mon, 3 day, 4 pre, 5 itr
    + indicates presence of .TMP output files, controls mnuview???.Enabled
* outselect(1 To 5)  As Integer
    + True(-1) if output selected for that time interval, False(0) if not
    + controls checking in Select Output menu
* strng$
* Title$(1 To gctMaxoutvars)
    + column titles for gctMaxoutvars variables
* txt$(1 To 30)
    + text strings for help
* userfont As String
* varno
    + output variable number, from 1 to gctMaxoutvars, always locally used
* yvars, mvars, dvars, pvars, ivars
    + number of vars/cols in output

Auxiliary numbers, constants, parameters and variables with an allocated value:

* dum         - dummy variable
* errnum      - error number from ERR
* evalnumber  - number of EVAL screens
* INIDAYS     - number of days to initialize run
* lstart      - starting output layer
* lend        - ending output layer
* lstep       - step for output layers
* parserr     - true if parse error in Soil parameters
* prfiletype  - precip file type, 1-csv,2-decimal period, 3-decimal comma
* rerunnumber - 0 for new run, 1 to 9 for each successive add run
* rstop       - if run should stop at end of day, 2 if EOF and date check on, 3 if crash error in run or immediate stop request, 4 if max number of graphs
* RUNDAYS     - number of output days to run after initialize
* runflag     - 0 for new run, 1 for add run, 2 for continue run

```{r}
ML <- 25    
gctMaxoutvars <- 60 
gctGenoutv <- 19 
gctIntoutv <- 22 
gctLayoutv <- 12 
gvals <- 8 
maxgraphs <- 100
#
#A <- numeric(ML) # parsed values from INSTRNG
#aparsed  # parsed values from gsbparsestring, first subscript is 0
#bad      # error indicator passed back to calling routine
#B90path$  # path for running program, used for .INI and .TMP
#cancelsave As Boolean #true if save or print should be cancelled
#chkout(1 To gctMaxoutvars, 1 To 5) As Integer #1 for output, 5 intervals
#        indicates which variables/interval selected
#        stored in .INI
#commadec As Boolean
#dfiletype    #data file type, 1-csv,2-decimal period, 3-decimal comma
#
dum   <-0  
errnum <-0   
#
#EV(1 To 25) As New frmevalout # EVAL windows
#
evalnumber <-0 
#
#FG(1 To maxgraphs) As New frmgraph  # multiple instances of graph
#gnumber   # graph number
#graphon    # true if graph has been initialized
#
INIDAYS<-0  
#
#inoutdir$  # latest input-output directory
#intrvl     #output interval index, 1 ANN, 2 MON, 3 DAY, 4 PRE, 5 ITR
#ivar(1 To gctMaxoutvars), pvar(1 To gctMaxoutvars), dvar(1 To gctMaxoutvars), mvar(1 To gctMaxoutvars), yvar(1 To gctMaxoutvars) # list of var numbers for output, does not count mulltiple soil layers
#
lstart <-0
lend <-0 
lstep <-0 
#
#msg As String#
#noenderror As Integer  # 0 if no END in parameter file, -1 if END is found
#noruncontinue #1 to prevent Run-Continue after soil parameter changes
#NLN As String  # new line
#outfilestatus(1 To 5) As Integer  #0 if no output, 1 if output, subscript 1 ann, 2 mon, 3 day, 4 pre, 5 itr
##        indicates presence of .TMP output files, controls mnuview???.Enabled
#outselect(1 To 5)  As Integer # True(-1) if output selected for that time interval, False(0) if not
#        controls checking in Select Output menu
#
parserr <-0  
prfiletype  <-0  
rerunnumber<-0 
rstop <-0 
RUNDAYS<-0  
runflag <-0
#
#strng$
#Title$(1 To gctMaxoutvars)   # column titles for gctMaxoutvars variables
#txt$(1 To 30) # text strings for help
#userfont As String
#varno # output variable number, from 1 to gctMaxoutvars, always locally used
#yvars, mvars, dvars, pvars, ivars  # number of vars/cols in output
```

### Constants
This table shows all constants which are used:

Constant|Value    |Description
--------|---------|-----------------------------------------------------------------------------------
DT      |1        |time step for DFILE interval,  must be 1 d
WTOMJ   |0.0864   |constant unit conversion, (MJ m-2 d-1)/(watt/m2) = 86400 s/d * .000001 MJ/J
ETOM    |0.4085   |(mm water)/(MJ/m2) using Lv 2448 MJ/Mg, density of water 1 Mg/m3 = 1E3 mm/m / (2448 MJ/Mg * 1 Mg/m3)
CPRHO   |1240     |volumetric heat capacity of air (J m-3 K-1)
GAMMA   |0.067    |psychrometer constant (kPa/K)
CVLQ    |0.00418  |volumetric heat capacity of water (MJ m-2 mm-1 K-1)
CVICE   |0.00192  |volumetric heat capacity of ice (MJ m-2 mm-1 K-1)
LF      |0.335    |heat of fusion of water (MJ m-2 mm-1)
LS      |2.824    |latent heat of sublimation of snow (MJ m-2 mm-1)
RHOWG   |0.00981  |density of water times gravity acceleration (MPa/m or kPa/mm)
SIGMA   |5.67E-8  |Stefan-Boltzmann constant (W m-2 K-4)
SC      |1367     |solar constant, value from [Lean (1991)](./literature.html),(W/m2)
K       |0.4      |vonKarman constant
PI      |3.1416   |Pi

```{r}
DT <- 1
WTOMJ <- 0.0864
ETOM <- 0.4085
CPRHO <- 1240
GAMMA <- 0.067
CVLQ <- 0.00418
CVICE <- 0.00192
LF <- 0.335
LS <- 2.824
RHOWG <- 0.00981
SIGMA <- 0.0000000567
SC <- 1367
K <- 0.4
PI <- 3.1416
```

### Parameters and variables
This list shows all variables that are used in alphabetic order:

* AA      - average available energy over daytime or nighttime (W/m2)
* ADEF    - available water deficit in root zone (mm), output only
* ALB     - * albedo with no snow
* ALBEDO  - albedo
* ALBSN   - * albedo with snow on the ground
* ALPHA   - modified Cowan alpha (MPa)
* ASPECT  - aspect, radians through east from north
* ASPECTD - * aspect, degrees through east from north
* ASUBS   - average avail. energy at ground over day or night (W/m2)
* ATR     - actual transpiration rate for daytime or night (mm/d)
* ATRANI  - actual transp.rate from layer for daytime or night (mm/d)
* ATRI    - actual transp.rate from layer for daytime and night (mm/d)
* AWAT    - available soil water in root zone (mm), output only

* BALERD  - error in water balance (mm)
* BALERM  - error in water balance (mm)
* BALERY  - error in water balance (mm)
* BEXP    - * exponent for psi-theta relation
* BYFL    - bypass flow rate from all layers for iteration (mm/d)
* BYFLI   - bypass flow rate from layer (mm/d)
* BYFLPI  - bypass flow from layer (mm)
* BYFLDI  - bypass flow from layer (mm)
* BYFLMI  - bypass flow from layer (mm)
* BYFLYI  - bypass flow from layer (mm)
* BYFLP   - total bypass flow (mm)
* BYFLD   - total bypass flow (mm)
* BYFLM   - total bypass flow (mm)
* BYFLY   - total bypass flow (mm)
* BYFRAC  - fraction of layer infiltration to bypass flow
* BYPAR   - * 1 to allow BYFL, or 0 to prevent BYFL

* C       - dummy input string
* C1      - * intercept of relation of solar rad. to sunshine duration
* C2      - * slope of relation of solar radiation to sunshine duration
* C3      - * longwave correction factor for overcast sky
* CC      - cold content of snowpack (positive) (MJ m-2)
* CCFAC   - * cold content factor (MJ m-2 d-1 K-1)
* CHM     - @ Clapp-Hornberger m parameter (kPa)
* CHN     - @ Clapp-Hornberger n parameter
* CINTRL  - * maximum interception storage of rain per unit LAI (mm)
* CINTRS  - * maximum interception storage of rain per unit SAI (mm)
* CINTSL  - * maximum interception storage of snow per unit LAI (mm)
* CINTSS  - * maximum interception storage of snow per unit SAI (mm)
* CR      - * light extinction coefficient for projected LAI + SAI
* CS      - * ratio of projected SAI to canopy height (m-1)
* CVPD    - * vapor pressure deficit at which conductance is halved (kPa)
* CZR     - * ratio of roughness to HEIGHT for rough closed canopies
* CZS     - * ratio of roughness to HEIGHT for smooth closed canopies

* DAYLEN  - daylength in fraction of day (d-1)
* DAYMO   - * days in month
* DD      - ** day of the month from data files
* DELTA   - dES/dT at some temperature T (kPa/K)
* DENSEF  - * density or thinning multiplier for MAXLAI,CS,RTLEN,RPLANT, not <.001
* DISP    - zero-plane displacement (m)
* DISPC   - zero-plane displacement for closed canopy of HEIGHT (m)
* DOM     - day of month
* DOY     - **** first day of the year from DFILE header
* DPSIDW  - rate of change of total potential with watercontent (kPa/mm)
* DPSIMX  - * maximum potential difference considered equal (kPa)
* DRAIN   - * multiplier of VFLUX(n) for drainage to groundwater
* DSFL    - downslope flow rate from all layers for iteration (mm/d)
* DSFLI   - downslope flow rate from layer (mm/d)
* DSFLP   - downslope flow (mm)
* DSFLD   - downslope flow (mm)
* DSFLM   - downslope flow (mm)
* DSFLY   - downslope flow (mm)
* DSFLPI  - downslope drainage from layer (mm)
* DSFLDI  - downslope drainage from layer (mm)
* DSFLMI  - downslope drainage from layer (mm)
* DSFLYI  - downslope drainage from layer (mm)
* DSLOPE  - slope for DSFL (radians)
* DSLOPED - * slope for DSFL (degrees)
* DSWMAX  - * maximum change allowed in SWATI (percent of SWATMX(i))
* DTI     - time step for iteration interval (d)
* DTIMAX  - * maximum iteration time step (d)
* DTINEW  - second estimate of DTI
* DTP     - @ time step for precipitation interval, may be <= 1 d
* DTRI    - time remaining in precipitation interval (d)
* DUMM    - dummy array for subroutine calls
* dummy   - dummy variable for subroutine calls
* DURATN  - * average duration of daily precip by month (hr)

* EA      - ** vapor pressure for the day (kPa)
* ES      - saturated vapor pressure (kPa)
* ESLOPE  - slope for evapotranspiration and snowmelt (radians)
* ESLOPED - * slope for evapotranspiration and snowmelt (degrees)
* EVAPP   - evapotranspiration (mm)
* EVAPD   - evapotranspiration (mm) 
* EVAPM   - evapotranspiration (mm) 
* EVAPY   - evapotranspiration (mm)

* FARR    - array of simulated daily flow for statistics
* FETCH   - * weather station fetch (m)
* FLOWP   - total flow (mm)
* FLOWD   - total flow (mm) 
* FLOWM   - total flow (mm)
* FLOWY   - total flow (mm)
* FRINTL  - * intercepted fraction of rain per unit LAI
* FRINTS  - * intercepted fraction of rain per unit SAI
* FSINTL  - * intercepted fraction of snow per unit LAI
* FSINTS  - * intercepted fraction of snow per unit SAI
* FXYLEM  - * fraction of plant resistance in xylem

* GER     - ground evaporation rate for daytime or night (mm/d)
* GEVP    - average ground evaporation for day (mm/d)
* GIR     - ground evap. rate with intercep. for daytime or night (mm/d)
* GIVP    - average ground evaporation for day with interception (mm/d)
* GLMAX   - maximum leaf conductance (m/s)
* GLMAXC  - * maximum leaf conductance (cm/s)
* GLMIN   - minimum leaf conductance (m/s)
* GLMINC  - * minimum leaf conductance (cm/s)
* GRAPH   - * runtime graphics output, 0-none, 1-continuous, 2-pause
* GRDMLT  - * rate of groundmelt of snowpack (mm/d)
* GSC     - * discharge from GWAT, fraction per day (d-1)
* GSP     - * fraction of discharge to seepage
* GWAT    - groundwater storage below soil layers (mm)
* GWATIN  - **** initial groundwater storage below soil layers (mm)
* GWFL    - streamflow rate from groundwater (mm/d)
* GWFLP   - groundwater flow (mm)
* GWFLD   - groundwater flow (mm)
* GWFLM   - groundwater flow (mm)
* GWFLY   - groundwater flow (mm)

* mnuhalfiter

* HEIGHT  - canopy height (m)
* HR      - * height above which CZR applies (m)
* HS      - * height below which CZS applies (m)

* I       - index variable for layer number
* I0HDAY  - potential insolation on horizontal (MJ m-2 d-1)
* IDAY    - day number in run
* II      - ** input precipitation interval number
* ILAYER  - number of layers over which infiltration is distributed
* IDEPTH  - * depth over which infiltration is distributed
* IMPERV  - * impervious fraction of area for SRFL
* INFEXP  - * infiltration exponent, 0-all to top to 1-uniform with depth
* INFRAC  - @ fraction of infiltration to each layer
* INFLI   - infiltration rate into layer (mm/d)
* INFLP   - infiltration into soil matrix (mm)
* INFLD   - infiltration into soil matrix (mm)
* INFLM   - infiltration into soil matrix (mm)
* INFLY   - infiltration into soil matrix (mm)
* INFLPI  - infiltration to layer (mm)
* INFLDI  - infiltration to layer (mm)
* INFLMI  - infiltration to layer (mm)
* INFLYI  - infiltration to layer (mm)
* INTR    - intercepted rain (mm)
* INTRIN  - & initial intercepted rain (mm)
* INTS    - intercepted snow (mm)
* INTSIN  - & initial intercepted snow (mm)
* IRVP    - evaporation rate of intercepted rain (mm/d)
* IRVPD   - evaporation of intercepted rain (mm)
* IRVPM   - evaporation of intercepted rain (mm)
* IRVPY   - evaporation of intercepted rain (mm)
* ISVP    - evaporation rate of intercepted snow (mm/d)
* ISVPD   - evaporation of intercepted snow (mm)
* ISVPM   - evaporation of intercepted snow (mm)
* ISVPY   - evaporation of intercepted snow (mm)

* J       - index variable for day-night separation

* KF      - * hydraulic conductivity at field capacity (mm/d)
* KK      - hydraulic conductivity (mm/d)
* KSAT    - @ saturated hydraulic conductivity (mm/d)
* KSNVP   - * multiplier to fix snow evaporation problem
* L1      - @ latitude of equivalent slope (radians)
* L2      - @ time shift of equivalent slope (radians)
* LAI     - leaf area index (m2/m2)
* LAIMLT  - * parameter for snowmelt dependence on LAI (Globalensionless)
* LAT     - latitude (radians)
* LATD    - **** latitude (degrees)
* LENGTH  - * slope length for DSFL (m)
* LPC     - * minimum LAI defining a closed canopy
* LWIDTH  - * leaf width (m)

* MARR    - array of measured daily flow for statistics
* MAXHT   - * maximum height for the year (m)
* MAXLAI  - * maximum projected leaf area index for the year (m2/m2)
* MAXLQF  - * maximum liquid water fraction of SNOW (Globalensionless)
* MELFAC  - * degree day melt factor for open (MJ m-2 d-1 K-1)
* MESFL   - ** measured streamflow for day (mm)
* MESFLD  - measured streamflow (mm)
* MESFLM  - measured streamflow (mm)
* MESFLY  - measured streamflow (mm)
* MESFLP  - ** measured streamflow rate for precip interval (mm/d)
* MM      - ** month from data files
* MONTHN  - month number
* MXKPL   - * maximum plant conductivity ((mm/d)/MPa)
* MXRTLN  - * maximum root length per unit land area (m/m2)

* N       - index variable for precipitation interval
* NN      - * wind/diffusivity extinction coefficient
* NDAYS   - * number of days in run
* NITS    - number of iterations in precipitation interval
* NITSD   - total number of iterations for day
* NITSM   - total number of iterations for month
* NITSY   - total number of iterations for year
* NITSR   - total number of iterations for run
* NLAYER  - * number of soil layers to be used in model, <= ML
* NOOUTF  - * 1 if no outflow allowed from roots, otherwise 0
* NTFLI   - net flow rate into layer (mm/d)
* NTFLPI  - net flow into layer (mm)
* NTFLDI  - net flow into layer (mm)
* NTFLMI  - net flow into layer (mm)
* NTFLYI  - net flow into layer (mm)

* PINT    - average potential interception for day (mm/d)
* PINTD   - potential interception (mm)
* PINTM   - potential interception (mm) 
* PINTY   - potential interception (mm)
* PIR     - potential interception rate for daytime or night (mm/d)
* PREC    - precipitation rate (mm/d)
* PRECD   - precipitation (mm)
* PRECM   - precipitation (mm)
* PRECY   - precipitation (mm)
* PREINT  - ** precipitation for precipitation interval (mm)
* PRECIN  - ** daily precipitation (mm)
* PSICR   - * minimum plant leaf water potential (MPa)
* PSIF    - * matric potential at field capacity (kPa)
* PSIG    - @ gravity potential (kPa)
* PSIM    - & matric soil water potential for layer (kPa)
* PSIMIN  - initial PSIM()
* PSITI   - total potential (kPa)
* PSNVP   - potential snow evaporation (mm/d)
* PTR     - potential transpiration rate for daytime or night (mm/d)
* PTRAN   - average potential transpiration rate for day (mm/d)
* PTRAND  - potential transpiration (mm)
* PTRANM  - potential transpiration (mm) 
* PTRANY  - potential transpiration (mm)

* QFFC    - * quick flow fraction (SRFL or BYFL) at field capacity
* QFPAR   - * quick flow parameter (SRFL or BYFL)
* QLAYER  - number of soil layers for SRFL, 0 to prevent SRFL
* QDEPTH  - * soil depth for SRFL calculation, 0 to prevent SRFL

* R5      - * solar radiation at which conductance is halved (W/m2)
* RAA     - Shuttleworth-Wallace atmosphere aerodynamic resistance (s/m)
* RAC     - Shuttleworth-Wallace canopy aerodynamic resistance (s/m)
* RAS     - Shuttleworth-Wallace ground aerodynamic resistance (s/m)
* RELHT   - * ten pairs of DOY and relative canopy height
* RELLAI  - * ten pairs of DOY and relative LAI
* ROOTDEN - * 25 pairs of root layer thickness (mm) and relative root density per unit volume
* RELDEN  - relative root density per unit volume for soil layers
* RFAL    - rainfall rate (mm/d)
* RFALD   - rainfall (mm)
* RFALM   - rainfall (mm)
* RFALY   - rainfall (mm)
* RHOTP   - * ratio of total leaf area to projected area
* RINT    - rainfall catch rate (mm/d)
* RINTD   - rain interception (mm)
* RINTM   - rain interception (mm)
* RINTY   - rain interception (mm)
* RM      - * maximum solar radiation, at which FR = 1 (W/m2)
* RNET    - rain reaching soil surface (mm/d)
* RNETD   - rainfall to soil surface (mm)
* RNETM   - rainfall to soil surface (mm)
* RNETY   - rainfall to soil surface (mm)
* RPLANT  - plant resistivity to water flow (MPa d/mm)
* RROOTI  - root resistance for layer (MPa d/mm)
* RTHR    - rain throughfall rate (mm/d)
* RTHRD   - rain throughfall (mm)
* RTHRM   - rain throughfall (mm)
* RTHRY   - rain throughfall (mm)
* RTLEN   - root length per unit land area (m/m2)
* RTRAD   - * average root radius (mm)
* RRD     - Default Radiation Ratio
* RSC     - Shuttleworth-Wallace canopy surface resistance (s/m)
* RSNO    - rain added to snowpack (mm/d)
* RSNOD   - rain on snow (mm)
* RSNOM   - rain on snow (mm)
* RSNOY   - rain on snow (mm)
* RSS     - Shuttleworth-Wallace soil surface resistance (s/m)
* RSSA    - * soil evaporation resistance at field capacity (s/m) 
* RSSB    - * exponent in relation of soil evap res to water potential
* RSTEMP  - * base temperature for snow-rain transition (degC)
* RXYLEM  - xylem resistance (MPa d/mm)

* SAFRAC  - source area fraction
* SAI     - stem area index (m2/m2)
* SAIMLT  - * parameter for snowmelt dependence on SAI (Globalensionless)
* SEEP    - deep seepage loss from groundwater (mm/d)
* SEEPP   - seepage loss (mm)
* SEEPD   - seepage loss (mm)
* SEEPM   - seepage loss (mm)
* SEEPY   - seepage loss (mm)
* SFAL    - snowfall rate (mm/d)
* SFALD   - snowfall (mm)
* SFALM   - snowfall (mm)
* SFALY   - snowfall (mm)
* SHEAT   - @ average soil heat flux for the day (W/m2, fixed at 0)
* SINT    - snowfall catch rate (mm/d)
* SINTD   - snow interception (mm) 
* SINTM   - snow interception (mm)
* SINTY   - snow interception (mm)
* SLFDAY  - ratio of potential insolation on slope to horizontal (map area)
* SLFL    - input rate to soil surface (mm/d)
* SLFLP   - input to soil surface (mm)
* SLFLD   - input to soil surface (mm)
* SLFLM   - input to soil surface (mm)
* SLFLY   - input to soil surface (mm)
* SLFLI   - macropore infiltration rate down from layer (mm/d)
* SLFLPI  - vertical macropore infiltration from layer (mm)
* SLFLDI  - vertical macropore infiltration from layer (mm)
* SLFLMI  - vertical macropore infiltration from layer (mm)
* SLFLYI  - vertical macropore infiltration from layer (mm)
* SLRAD   - average solar radiation on slope over daytime (W/m2)
* SLVP    - evaporation rate from soil (mm/d)
* SLVPD   - soil evaporation (mm)
* SLVPM   - soil evaporation (mm)
* SLVPY   - soil evaporation (mm)
* SMLT    - melt drainage rate from snowpack (mm/d)
* SMLTD   - snowmelt (mm)
* SMLTM   - snowmelt (mm)
* SMLTY   - snowmelt (mm)
* SNODEN  - * snow density (mm/mm)
* SNOEN   - energy flux density to snow surface (MJ m-2 mm-1 d-1)
* SNOFRC  - fraction of precipitation for the day as snow (unitless)
* SNOW    - water equivalent of snow on the ground (mm)
* SNOWIN  - **** initial water equivalent of snow on the ground (mm)
* SNOWLQ  - liquid water content of snow on the ground (mm)
* SNVP    - evaporation rate from snowpack (mm/d)
* SNVPD   - evaporation from snowpack (mm)
* SNVPM   - evaporation from snowpack (mm)
* SNVPY   - evaporation from snowpack (mm)
* SOLRAD  - ** solar radiation for the day, horizontal surface (MJ/m2)
* SOLRADC - SOLRAD as corrected if necessary by WEATHER routine (MJ/m2)
* SRFL    - source area flow rate (mm/d)
* SRFLP   - source area flow (mm)
* SRFLD   - source area flow (mm)
* SRFLM   - source area flow (mm)
* SRFLY   - source area flow (mm)
* STHR    - snow throughfall rate (mm/d)
* STHRD   - snow throughfall (mm)
* STHRM   - snow throughfall (mm)
* STHRY   - snow throughfall (mm)
* STONEF  - * stone volume fraction (unitless)
* STORD   - total water storage in system (mm)
* STORM   - total water storage in system (mm)
* STORY   - total water storage in system (mm)
* STRES   - TRAN / PTRAN for time period
* STRX    - string variable to trap q or esc
* SWAT    - total soil water in all layers (mm)
* SWATI   - water volume in layer (mm)
* SWATMX  - maximum water storage for layer (mm)
* SWATQF  - @water storage at field capacity for layers 1 to QLAYER (mm)
* SWATQX  - @ maximum water storage for layers 1 to QLAYER (mm)

* T1      - * lowest temp. at which stomates not temp. limited (degC)
* T2      - * highest temp. at which stomates not temp. limited (degC)
* TA      - mean temperature for the day at reference height (degC)
* TADTM   - average daytime temperature at reference height (degC)
* TAJ     - TADTM or TANTM depending on J
* TANTM   - average nighttime temperature at reference height (degC)
* TEMP    - temporary integer variable - apparently no longer used
* TH      - * temperature above which stomates are closed (degC)
* THETA   - water content (mm water / mm soil matrix)
* THETAF  - * volumetric water content at field capacity
* THICK   - * layer thicknesses (mm)
* THSAT   - * theta at saturation, matrix porosity
* TL      - * temperature below which stomates are closed (degC)
* TMAX    - ** maximum temperature for the day (degC)
* TMIN    - ** minimum temperature for the day (degC)
* TRANI   - average transpiration rate from layer (mm/d)
* TRANP   - transpiration (mm)
* TRAND   - transpiration (mm)
* TRANM   - transpiration (mm)
* TRANY   - transpiration (mm)
* TRANPI  - layer transpiration (mm)
* TRANDI  - layer transpiration (mm)
* TRANMI  - layer transpiration (mm)
* TRANYI  - layer transpiration (mm)
* TSNOW   - snowpack temperature (isothermal assumed, degC)

* UA      - average wind speed for the day at reference height (m/s)
* UADTM   - average wind speed for daytime at reference height (m/s)
* UAJ     - UADTN or UANTM depending on J
* UANTM   - average wind speed for nighttime at reference height (m/s)
* UW      - ** average wind speed for day at weather station (m/s)

* VPD     - vapor pressure deficit at reference height (kPa)
* VRFLI   - vertical matrix drainage rate from layer (mm/d)
* VRFLPI  - vertical matrix drainage from layer (mm)
* VRFLDI  - vertical matrix drainage from layer (mm)
* VRFLMI  - vertical matrix drainage from layer (mm)
* VRFLYI  - vertical matrix drainage from layer (mm)
* VV      - temporary VRFLI

* WETC    - @ wetness at PSICR (Globalensionless)
* WETF    - @ wetness at field capacity (Globalensionless)
* WETFR   - fraction of precipitation interval that canopy is wet
* WETINF  - * wetness at dry end of near-saturation range
* WETNES  - wetness (fraction of saturation)
* WNDRAT  - * ratio of nighttime to daytime wind speed

* XMAX    - maximum value for x-axis
* XMIN    - minimum value for x-axis

* YEARN   - **** first year from DFILE header
* YMAX    - * maximum value for y-axis
* YMIN    - @ minimum value for y-axis
* YNAME   - @ variable name
* YVAL    - plot value for y-axis
* YY      - ** year from data files, used only to check dates and to determine end of year for output

* Z0      - roughness parameter (m)
* Z0C     - roughness parameter for closed canopy of HEIGHT (m)
* Z0G     - * ground surface roughness (m)
* Z0GS    - ground or snow surface roughness (m)
* Z0S     - * snow surface roughness (m)
* Z0W     - * weather station roughness parameter (m)
* ZA      - reference height for TA, EA, UA, above ground (m)
* ZMINH   - * ZA minus HEIGHT, reference height above canopy top (m)
* ZW      - * weather station measurement height for wind (m)

```{r}
AA  <-0       
ADEF<-0   
ALB   <-0.07       
ALBEDO<-0    
ALBSN <-0.3           
ALPHA<-numeric(ML) 
ASPECT  <-0         
ASPECTD <-0       
ASUBS   <-0          
ATR<-numeric(2)          
ATRANI<-numeric(ML)  
ATRI<-matrix(0,2,ML) 
AWAT    <-0          
BALERD<-0 
BALERM<-0  
BALERY<-0 
BEXP<-rep(0,ML)  
BEXP[1]<-5.37433
BEXP[2]<-4.03320
BEXP[3]<-5.64096
BEXP[4]<-6
BEXP[5]<-5
BYFL   <-0       
BYFLI<-numeric(ML)  
BYFLPI<-numeric(ML)
BYFLDI<-numeric(ML)
BYFLMI<-numeric(ML)
BYFLYI<-numeric(ML)
BYFLP<-rep(0,ML)
BYFLD<-rep(0,ML) 
BYFLM<-rep(0,ML)  
BYFLY<-rep(0,ML)
BYFRAC<-numeric(ML)  
BYPAR<-1            
#C    <-0             
C1    <-0.25
C2    <-0.5     
C3    <-0.2            
CC      <-0   
CCFAC  <-0.3          
CHM<-numeric(ML)     
CHN<-numeric(ML)     
CINTRL<-0.15            
CINTRS<-0.15            
CINTSL<-0.6         
CINTSS<-0.6             
CR   <-0.5              
CS  <-0.035           
CVPD <-2           
CZR  <-0.05              
CZS <-0.13          
DAYLEN <-0           
DAYMO<-numeric(12)     
DD  <-0            
DELTA <-0           
DENSEF <-0.8580571           
DISP  <-0        
DISPC  <-0           
DOM  <-0        
DOY  <-0            
DPSIDW<-numeric(ML) 
DPSIMX  <-0.01          
DRAIN  <-0.626259     
DSFL  <-0         
DSFLI<-numeric(ML)   
DSFLP<-0 
DSFLD<-0 
DSFLM<-0 
DSFLY<-0   
DSFLPI<-numeric(ML)
DSFLDI<-numeric(ML)
DSFLMI<-numeric(ML)
DSFLYI<-numeric(ML)
DSLOPE  <-0         
DSLOPED  <-0      
DSWMAX   <-2        
DTI   <-0          
DTIMAX  <-0.5          
DTINEW  <-0.0          
DTP    <-1           
DTRI  <-0             
DUMM<-rep(0,gctMaxoutvars)  
dummy   <-0          
DURATN<-c(4,4,4,4,4,4,4,4,4,4,4,4)   
EA    <-0             
ES   <-0              
ESLOPE  <-0           
ESLOPED  <-2         
EVAPP<-0 
EVAPD<-0  
EVAPM<-0  
EVAPY <-0
FARR<-rep(0,366)    
FETCH   <-5000           
FLOWP<-0 
FLOWD<-0 
FLOWM<-0 
FLOWY<-0  
FRINTL  <-0.06           
FRINTS  <-0.06        
FSINTL  <-0.04           
FSINTS  <-0.04          
FXYLEM  <-0.5           
GER<-numeric(2)            
GEVP  <-0       
GIR<-numeric(2)          
GIVP <-0       
GLMAX  <-0           
GLMAXC  <-0.494377          
GLMIN <-0           
GLMINC  <-0.03        
GRAPH <-0            
GRDMLT <-0.35           
GSC   <-0            
GSP  <-0.085              
GWAT   <-0            
GWATIN <-20            
GWFL  <-0             
GWFLP<-0 
GWFLD<-0 
GWFLM<-0 
GWFLY<-0   
mnuhalfiter<-FALSE
HEIGHT  <-0          
HR   <-10              
HS    <-1            
#I    <-0            
I0HDAY <-0            
IDAY<-0             
II <-0               
ILAYER <-0          
IDEPTH <-1000            
IMPERV  <-0.025 
INFEXP  <-0.8797636          
INFRAC<-rep(0,ML)  
INFLI<-rep(0,ML) 
INFLP<-rep(0,ML)
INFLD<-rep(0,ML)
INFLM<-rep(0,ML)
INFLY<-rep(0,ML) 
INFLPI<-rep(0,ML)
INFLDI<-rep(0,ML)
INFLMI<-rep(0,ML)
INFLYI<-rep(0,ML)
INTR   <-0           
INTRIN  <-0           
INTS   <-0            
INTSIN <-0            
IRVP <-0              
IRVPD<-0 
IRVPM<-0 
IRVPY <-0             
ISVP  <-0           
ISVPD<-0 
ISVPM<-0 
ISVPY<-0  
J   <-0              
KF<-rep(0,ML)      
KF[1]<-6.9
KF[2]<-2.7
KF[3]<-2.9
KF[4]<-1
KF[5]<-3.5
KK<-rep(0,ML)      
KSAT<-rep(0,ML)    
KSNVP <-0.009734            
L1  <-0               
L2  <-0               
LAI  <-0              
LAIMLT <-0.2           
LAT   <-0             
LATD <-50.5              
LENGTH  <-0           
LPC  <-4              
LWIDTH    <-0.004      
MARR<-c(seq(1,366,1))    
MAXHT<-25             
MAXLAI  <-7.693270           
MAXLQF <-0.05            
MELFAC <-1.728930            
MESFL <-0            
MESFLD<-0
MESFLM<-0
MESFLY<-0 
MESFLP <-0           
MM    <-0     
MONTHN <-0        
MXKPL  <-7.03463      
MXRTLN <-3000.001    
N   <-0            
NN  <-2.5             
NDAYS<-0            
NITS<-0             
NITSD<-0            
NITSM<-0           
NITSY<-0            
NITSR<-0            
NLAYER  <-5         
NOOUTF <-1          
NTFLI<-rep(0,ML)  
NTFLPI<-rep(0,ML)
NTFLDI<-rep(0,ML)
NTFLMI<-rep(0,ML)
NTFLYI<-rep(0,ML)
PINT <-0        
PINTD<-0 
PINTM<-0 
PINTY <-0 
PIR<-c(0,0)          
PREC  <-0      
PRECD<-0 
PRECM<-0 
PRECY<-0             
PREINT  <-0          
PRECIN  <-0          
PSICR  <--2           
PSIF<-rep(0,ML)   
PSIF[1]<--11.818
PSIF[2]<--11.516 
PSIF[3]<--10.22
PSIF[4]<--10
PSIF[5]<--10
PSIG<-rep(0,ML)     
PSIM<-rep(0,ML)      
PSIMIN<-rep(-10,ML)  
PSITI<-rep(0,ML)    
PSNVP <-0          
PTR<-c(0,0)          
PTRAN  <-0          
PTRAND<-0 
PTRANM<-0 
PTRANY<-0          
QFFC<-0.00104        
QFPAR <-0.834524        
QLAYER <-1        
QDEPTH <-0.1          
R5    <-100           
RAA    <-0           
RAC   <-0            
RAS  <-0           
RELHT<-c(1,1,366,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)    
RELLAI<-c(1,1,366,1,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)   
ROOTDEN<-c(50,1,50,1,50,1,50,1,50,1,50,.3,50,.2,50,.1,50,.1,50,.1,50,0.0,50,0.0,50,0,50,0,50,0,50,0,50,0,50,0,50,0,50,0,50,0,50,0,50,0,50,0,50,0)  
RELDEN<-rep(0,ML)  
RFAL   <-0          
RFALD<-0
RFALM<-0
RFALY <-0            
RHOTP <-2            
RINT  <-0           
RINTD<-0
RINTM<-0
RINTY<-0           
RM <-1000               
RNET   <-0          
RNETD<-0
RNETM<-0 
RNETY <-0            
RPLANT<-0            
RROOTI<-rep(0,ML)    
RTHR   <-0           
RTHRD<-0
RTHRM<-0
RTHRY<-0            
RTLEN <-0           
RTRAD <-0.35           
RRD<-0.55           
RSC <-0              
RSNO <-0           
RSNOD<-0
RSNOM<-0
RSNOY<-0         
RSS<-0              
RSSA <-500             
RSSB <-1             
RSTEMP<- -1.29978         
RXYLEM <-0           
SAFRAC <-0           
SAI <-0              
SAIMLT <-0.5           
SEEP <-0             
SEEPP<-0
SEEPD<-0
SEEPM<-0
SEEPY<-0             
SFAL <-0     
SFALD<-0
SFALM<-0
SFALY<-0             
SHEAT<-0        
SINT <-0           
SINTD<-0 
SINTM<-0
SINTY <-0
SLFDAY  <-0          
SLFL  <-0            
SLFLP<-0
SLFLD<-0
SLFLM<-0
SLFLY<-0
SLFLI<-rep(0,ML)  
SLFLPI<-rep(0,ML)
SLFLDI<-rep(0,ML)
SLFLMI<-rep(0,ML)
SLFLYI<-rep(0,ML)
SLRAD <-0            
SLRADd<-0
SLVP <-0            
SLVPD<-0
SLVPM<-0
SLVPY<-0          
SMLT  <-0            
SMLTD<-0
SMLTM<-0 
SMLTY <-0           
SNODEN <-0.3         
SNOEN <-0           
SNOFRC <-0        
SNOW <-0             
SNOWIN  <-20          
SNOWLQ<-0            
SNVP  <-0            
SNVPD<-0
SNVPM<-0
SNVPY <-0            
SOLRAD <-0           
SOLRADC<-0           
SRFL <-0             
SRFLP<-0
SRFLD<-0
SRFLM<-0
SRFLY <-0
STHR <-0             
STHRD<-0
STHRM<-0
STHRY<-0 
STONEF<-rep(0.00,ML)
STONEF[1]<-0.02
STONEF[2]<-0.2
STONEF[3]<-0.25
STONEF[4]<-0.25
STONEF[5]<-0.7
STORD<-0
STORM<-0
STORY<-0 
STRES<-0            
STRX<-0             
SWAT<-0    
SWATI<-rep(0,ML)  
SWATMX<-rep(0,ML)  
SWATQF   <-0         
SWATQX <-0           
T1    <-10           
T2    <-30           
TA   <-40            
TADTM <-0           
TAJ  <-0             
TANTM <-0            
TEMP  <-0           
TH   <-40            
THETA<-rep(0,ML)   
THETAF<-rep(0,ML) 
THETAF[1]<-0.34062341
THETAF[2]<-0.39705807  
THETAF[3]<-0.24359704
THETAF[4]<-0.35
THETAF[5]<-0.23
THICK<-rep(0,ML)   
THICK[1]<-50
THICK[2]<-250
THICK[3]<-300
THICK[4]<-300
THICK[5]<-100
THSAT<-rep(0,ML)  
THSAT[1]<-0.6313342
THSAT[2]<-0.6189386
THSAT[3]<-0.4930716
THSAT[4]<-0.680
THSAT[5]<-0.600
TL    <-0           
TMAX  <-0            
TMIN <-0             
TRANI<-rep(0,ML)   
TRANP<-0
TRAND<-0
TRANM<-0
TRANY<-0             
TRANPI<-rep(0,ML)
TRANDI<-rep(0,ML)
TRANMI<-rep(0,ML) 
TRANYI<-rep(0,ML)
TSNOW  <-0          
UA     <-0           
UADTM  <-0           
UAJ   <-0           
UANTM <-0          
UW   <-0             
VPD  <-0             
VRFLI<-rep(0,ML)  
VRFLPI<-rep(0,ML)
VRFLDI<-rep(0,ML)
VRFLMI<-rep(0,ML)
VRFLYI<-rep(0,ML)
VV<-rep(0,ML)      
WETC<-rep(0,ML)    
WETF<-rep(0,ML)   
WETFR  <-0         
WETINF<-rep(0,ML)  
WETINF[1]<-0.92
WETINF[2]<-0.92
WETINF[3]<-0.92
WETINF[4]<-0.92
WETINF[5]<-0.92
WETNES<-rep(0,ML)  
WNDRAT <-.3           
XMAX  <-0            
XMIN  <-0           
YEARN <-0           
YMAX<-numeric(gvals)      
YMIN<-numeric(gvals)      
YNAME<-numeric(gvals)    
YVAL<-numeric(gvals)      
YY   <-0            
Z0   <-0             
Z0C  <-0             
Z0G  <-0.02             
Z0GS  <-0            
Z0S  <-0.001             
Z0W  <-2.8             
ZA   <-0            
ZMINH <-2           
ZW  <-14              
```