---
title: "B90V4"
author: "Rico Kronenberg, Lisa Marie Oehlschlägel"
date: "11 Mai 2018"
output: html_document
---

### B90V4 - BROOK90 Version 4
With running the MainProg.R script all necessary scripts and variables are loaded. Starting the model requires to run the script B90V4.R from source.

This Code is modified for Version 4, June 3, 1999. The Symbol ^^ shows variables that are returned from or altered by subroutines (requires a fixed width font). There are intrinsic functions needed:

* CSNG
* INT
* INKEY$
* CHR$

### Contents

* [Programm initialisations](#programm-initialisations)
* [Parameter initializations](#paramter-initialisations)
* [Running the model](#running-the-model)
* [Plotted data](#plotted-data)

### Programm initialisations
First the script of the main programm [MainProg.Rmd](./MainProg.Rmd) has to be rendered to run the script of B90V4.Rmd in R-Markdown.

```{r include=FALSE}
rmarkdown::render('MainProg.Rmd')
```

The program initializations starts if it is New Run or Rerun. Settings for the time are done, like number of days per month, interval times and zeroed vectors are set. The Year dates are changed from two-digit year to four-digit. The DOY is set with the help of [Function DOYF](./B90V4_sub.Rmd) with the input DOM, MONTHN and DAYMO and number of days for february depending on leap year or not. The time step for precipitation interval is calculated and zero accumulators are called. Some neccesary values are calculated or unlisted from functions:

* initial values
* soil water parameters and initial variables
* initial soil water variables
* initial total water in system
* any initial snow has zero liquid water and cold content

```{r}
if((runflag == 0) || (runflag == 1)){
  DAYMO[1] = 31
  DAYMO[2] = 28
  DAYMO[3] = 31
  DAYMO[4] = 30 
  DAYMO[5] = 31
  DAYMO[6] = 30
  DAYMO[7] = 31
  DAYMO[8] = 31
  DAYMO[9] = 30
  DAYMO[10] = 31
  DAYMO[11] = 30
  DAYMO[12] = 31
	IDAY =1
	IInterValDay=1
	NDAYS=length(MData[[1]])
	NITSR = 0
	NITSY = 0
	NITSM = 0
	YEARN = as.numeric(MData[[1]][IDAY])
	
	daymax=NDAYS-IDAY+1
	maxF=0
	precc=rep(0,daymax)
	evpp=rep(0,daymax)
	floww=rep(0,daymax)
	rnett=rep(0,daymax)
	ptrann=rep(0,daymax)
	irvpp=rep(0,daymax)
	isvpp=rep(0,daymax)
	snoww=rep(0,daymax)
	swatt=rep(0,daymax)
	pintt=rep(0,daymax)
	snvpp=rep(0,daymax)
	slvpp=rep(0,daymax)
	trandd=rep(0,daymax)
	mesfld=rep(0,daymax)
	smltd=rep(0,daymax)
	slfld=rep(0,daymax)
	rfald=rep(0,daymax)
	sfald=rep(0,daymax)
	awatt=rep(0,daymax)
	adeff=rep(0,daymax)
	sintdd=rep(0,daymax)
	rintdd=rep(0,daymax)
	rthrd=rep(0,daymax)
	sthrd=rep(0,daymax)
	rsnod=rep(0,daymax)
	
	if( YEARN < 100){
		if(YEARN > 20){
			YEARN = YEARN + 1900
		}else{
			YEARN = YEARN + 2000
		}
	}
  MONTHN = as.numeric(MData[[2]][IDAY])
	DOM = as.numeric(MData[[3]][IDAY])
	DOY = DOY=DOYF(DOM,MONTHN,DAYMO)
		
	  if (fnleap()) {
			DAYMO[2] = 29
		}else{
			DAYMO[2] = 28
		}
	if (SUBDAYDATA) {
  	DTP = DT / NPINT
	}else{
		DTP = DT
	}
# zero accumulators
	zyear()
	zmonth()
# initial values
	SNOW = SNOWIN
	GWAT = GWATIN
	INTR = INTRIN
	INTS = INTSIN
	for( i in 1:NLAYER){
		PSIM[i] = PSIMIN[i]
	}
# soil water parameters and initial variables
	soilp<-SOILPAR()
	  PSIG<-unlist(soilp[2])
	  SWATMX<-unlist(soilp[3])
	  WETF<-unlist(soilp[4])
	  WETC<-unlist(soilp[5])
	  CHM<-unlist(soilp[6])
	  CHN<-unlist(soilp[7]) 
	  WETNES<-unlist(soilp[8])
	  SWATI<-unlist(soilp[9])
	  KSAT<-unlist(soilp[10])
# ^^
# initial soil water variables
	soil<-SOILVAR()
	  PSITI<-soil[1:ML]
	  THETA<-soil[(ML+1):(2*ML)]
	  KK<-soil[(2*ML+1):(3*ML)]
	  SWAT<-soil[(3*ML+1)]
# ^^
# initial total water in system
	STORD = INTR + INTS + SNOW + SWAT + GWAT
	STORM = STORD
	STORY = STORD
# any initial snow has zero liquid water and cold content
	CC = 0
	SNOWLQ = 0
}
```

### Parameter initializations
Parameter initializations are done for New Run, Rerun and Continue Run. First some parameters are converted. Then the equivalent slope for radiation calculations is extracted with the help of [Function EQUIVSLP](./SUN.Rmd). Other initialized parameters are:

* infiltration parameters
* source area parameters
* root density parameters

```{r}
# parameter conversions
GLMAX = GLMAXC / 100
GLMIN = GLMINC / 100
LAT = LATD / 57.296
ESLOPE = ESLOPED / 57.296
DSLOPE = DSLOPED / 57.296
ASPECT = ASPECTD / 57.296
# equivalent slope for radiation calculations
equi<-EQUIVSLP(LAT, ESLOPE, ASPECT)
  L1<-unlist(equi[1])
  L2<-unlist(equi[2])
# ^^
# infiltration parameters
infpa<-INFPAR(INFEXP, IDEPTH, NLAYER, THICK)
  ILAYER<-unlist(infpa[1])
  INFRAC<-unlist(infpa[2])
# ^^
# source area parameters
srfp<-SRFPAR(QDEPTH, NLAYER, THETAF, THICK, STONEF, SWATMX)
  QLAYER<-unlist(srfp[1]) 
  SWATQX<-unlist(srfp[2])
  SWATQF<-unlist(srfp[3])
# ^^
# root density parameters
RELDEN<-RTDEN(ROOTDEN, NLAYER, THICK)
```

### Running the model
After all initialization steps are done the model can be started. All important steps that are done are written in the following script as comments.

```{r}
while( IDAY <= NDAYS){  
	NITSD = 0
		subdatafileline(IDAY)
  if( IDAY == INIDAYS + 1){
# end of initialization, reinitialize year and month accumulators
		STORD = INTR + INTS + SNOW + SWAT + GWAT
		STORM = STORD
		STORY = STORD
		NITSY = 0
		NITSM = 0
		zyear()
		zmonth()
  }
# calculate derived variables
	MSBSETVARS()
#
#* * * * *  B E G I N   D A Y - N I G H T   E T   L O O P  * * * * * * * * *
#potential and actual interception, evaporation, and transpiration
	MSBDAYNIGHT()
#
#* * * * * * * *  E N D   D A Y - N I G H T   L O O P  * * * * * * * * * *
# average rates over day
	PTRAN = (PTR[1] * DAYLEN + PTR[2] * (1 - DAYLEN)) / DT
	GEVP = (GER[1] * DAYLEN + GER[2] * (1 - DAYLEN)) / DT
	PINT = (PIR[1] * DAYLEN + PIR[2] * (1 - DAYLEN)) / DT
	GIVP = (GIR[1] * DAYLEN + GIR[2] * (1 - DAYLEN)) / DT
	for(i in 1:NLAYER){
		TRANI[i] = (ATRI[1, i] * DAYLEN + ATRI[2, i] * (1 - DAYLEN)) / DT
	}
# zero daily integrators
	zday()
#
#* * * * * * * * B E G I N   P R E C I P   I N T E R V A L * * * * * * * * *
for( N in 1:NPINT){  
	if (SUBDAYDATA){
    subprfileline(IInterValDay)
	  if (MESFLP <= -0.01) {MESFLP = MESFL / DT}
	}else{
# precip data from data file
		PREINT = PRECIN / DT
		MESFLP = MESFL / DT
	}
# interception and snow accumulation/melt
	MSBPREINT()
# initialize for iterations
# initial time remaining in iteration time step = precip time step
	DTRI = DTP
# initialize iteration counter
	NITS = 0
# zero precip interval integrators
	zpint()
#
#  *  *  *  *  *  *  B E G I N   I T E R A T I O N   *  *  *  *  *  *  *  *
while(!(DTRI <= 0)){  
		NITS = NITS + 1
# check for events
		if (NITS %% 100 == 0) {}
# water movement through soil
		MSBITERATE() 
# iteration calculations
# calculate SLFLI vertical macropore infiltration out of layer
		SLFLI[1] = SLFL - INFLI[1] - BYFLI[1]
		if (ILAYER >= 2){
		  if (NLAYER >= ILAYER +1){
		    for (i in 2:ILAYER){ 
		    # does not execute if ILAYER% = 1 or 0
			    SLFLI[i] = SLFLI[i - 1] - INFLI[i] - BYFLI[i]
		    }
		    for( i in (ILAYER + 1):NLAYER){ 
		    # does not execute if NLAYER% < ILAYER% + 1
			    SLFLI[i] = 0
		    }
		  }
		}
# integrate below ground storages over iteration interval
		for( i in 1:NLAYER){
			SWATI[i] = SWATI[i] + NTFLI[i] * DTI
		}
		GWAT = GWAT + (VRFLI[NLAYER] - GWFL - SEEP) * DTI
# new soil water variables and test for errors
		for (i in 1:NLAYER){
			swchek(i)
			WETNES[i] = SWATI[i] / SWATMX[i]
			PSIM[i] = FPSIMF(WETNES[i], PSIF[i], BEXP[i], WETINF[i], WETF[i], CHM[i], CHN[i])
		}
		soil<-SOILVAR()
		   PSITI<-soil[1:ML]
		   THETA<-soil[(ML+1):(2*ML)]
		   KK<-soil[(2*ML+1):(3*ML)]
		   SWAT<-soil[(3*ML+1)]
# ^^
# iteration output
# flows accumulated over precip interval
	paccum()
# time remaining in precipitation time-step
	DTRI = DTRI - DTI
	NITSR = NITSR + 1  # for visible display of iterations
}
#
#  *  *  *  *   E N D   i T E R A T I O N    L O O P  *  *  *  *  *  *  *  *
# display iterations
# integrate interception storages over precip interval
INTS = INTS + (SINT - ISVP) * DTP
INTR = INTR + (RINT - IRVP) * DTP
#  flows for precip interval summed from components
psum()
# precipitation interval output
# flows accumulated over day
daccum()
# accumulate iterations
  NITSD = NITSD + NITS
  NITSM = NITSM + NITS
  NITSY = NITSY + NITS
  IInterValDay<-IInterValDay+1
}
#
#* * * * *  E N D   P R E C I P   I N T E R V A L   L O O P  * * * * * * * *
# flows for day summed from components
dsum()
# check for water balance error
BALERD = STORD - (INTR + INTS + SNOW + SWAT + GWAT) + PRECD - EVAPD - FLOWD - SEEPD
STORD = INTR + INTS + SNOW + SWAT + GWAT
# flows accumulated over month
maccum()
# date checking on
if(DOM == DAYMO[MONTHN]){
# set up for next month
zmonth()
MONTHN = MONTHN + 1
DOM = 0
NITSM = 0
}  # for end of month
if (MONTHN == 13) {
# end of year
# set up for next year
  MONTHN = 1
  DOM = 1
  DOY = 1
  YEARN = YEARN + 1
  zyear()
  if (fnleap() ){
    DAYMO[2] = 29
  }else{
    DAYMO[2] = 28
  }
NITSY = 0
NITSM = 0
} 
#set up for next day
IDAY = IDAY + 1
  MONTHN = as.numeric(MData[[2]][IDAY])
  DOM = as.numeric(MData[[3]][IDAY])
  YEARN = as.numeric(MData[[1]][IDAY])
  if(IDAY <= NDAYS)
  DOY=DOYF(DOM,MONTHN,DAYMO)
 
#* * * I N P U T   W E A T H E R   L I N E   F R O M   D F I L E * * *
#subdatafileline()
#
# ***************   E N D    D A Y   L O O P    **************************
	precc[daymax-NDAYS+IDAY-1]<-PRECD
	evpp[daymax-NDAYS+IDAY-1]<-EVAPD
	floww[daymax-NDAYS+IDAY-1]<-FLOWD
	rnett[daymax-NDAYS+IDAY-1]<-RNET
	irvpp[daymax-NDAYS+IDAY-1]<-IRVPD
	isvpp[daymax-NDAYS+IDAY-1]<-ISVPD
	ptrann[daymax-NDAYS+IDAY-1]<-PTRAND
	snoww[daymax-NDAYS+IDAY-1]<-SNOW
	swatt[daymax-NDAYS+IDAY-1]<-SWAT
	pintt[daymax-NDAYS+IDAY-1]<-PINTD
	snvpp[daymax-NDAYS+IDAY-1]<-SNVPD 
	slvpp[daymax-NDAYS+IDAY-1]<-SLVPD 
	trandd[daymax-NDAYS+IDAY-1]<-TRAND
	mesfld[daymax-NDAYS+IDAY-1]<-MESFLD
	smltd[daymax-NDAYS+IDAY-1]<-SMLTD
	slfld[daymax-NDAYS+IDAY-1]<-SLFLD
	rfald[daymax-NDAYS+IDAY-1]<-RFALD
	awatt[daymax-NDAYS+IDAY-1]<-AWAT
	adeff[daymax-NDAYS+IDAY-1]<-ADEF
	sintdd[daymax-NDAYS+IDAY-1]<-SINTD
	rintdd[daymax-NDAYS+IDAY-1]<-RINTD
	sfald[daymax-NDAYS+IDAY-1]<-SFALD
	rthrd[daymax-NDAYS+IDAY-1]<-RTHRD
	sthrd[daymax-NDAYS+IDAY-1]<-STHRD
	rsnod[daymax-NDAYS+IDAY-1]<-RSNOD
}
```

### Plotted data
At the end some data are plotted:

* precipitation - precc (points)
* evpp          - evaporation (blue line)
* mesfld        - measured flow (darkgreen line)
* floww         - total flow (darkgreen line)

```{r}
plot(1:NDAYS,floww[1:NDAYS],col="red",type="l",lwd=3,ylim=c(0.0, 70),xlab="Tage [d]",ylab="Werte [mm/d]")

PrecBound<-10
points( which(precc %in% precc[precc>PrecBound]),precc[precc>PrecBound],col="blue",pch="*")
lines(1:NDAYS,evpp,col="blue",lwd=3)
lines(1:NDAYS,mesfld,col="red",lwd=2)
lines(1:NDAYS,floww,col="darkgreen",lwd=3)
```

![alt text][output]

[output]: C:/Users/LM/Documents/BROOK90/Documentation/RMD_Files/output.png "Logo Title Text 2"

