---
title: "SUN"
author: "Rico Kronenberg, Lisa Marie Oehlschlägel"
date: "17 April 2018"
output: html_document
---

### SUN - Radiation
Daily solar radiation on a horizontal surface (SOLRAD in MJ/m2) is an input variable to BROOK90. This value is sometimes called global radiation to emphasize that it includes both direct or beam radiation from the sun and diffuse radiation from the sky hemisphere. It is directly measured by various types of pyranometers at most research sites and some locations of the National Weather Service. SOLRAD is internally prevented from being larger than 0.99 times the potential insolation (in the absence of an atmosphere), I0HDAY.

If no data are available and input SOLRAD is zero, BROOK90 uses a fixed fraction of the potential solar radiation on a horizontal surface. This fraction had been set at 0.55, but as of Version 4.8 it can be changed on the BROOK90 main window. Use of this generalized fraction loses the effects of day-to-day variation of solar radiation in the model. 

### Contents

* [Function AVAILEN](#function-availen)
* [Function EQUIVSLP](#function-equivslp)
* [Function SUNDS](#function-sunds)
* [Function FUNC3](#function-func3)
* [Function HAFDAY](#function-hafday)

### Function AVAILEN
AVAILEN calculates net radiation minus soil heat flux at the top and the bottom of the canopy, separately for daytime and for nighttime. 

Estimates of available energy above and below the canopy are made in subroutine AVAILEN. Available energy is net radiation minus subsurface heat flux (SHEAT), and is the energy available for partitioning into heating the air and evaporating water. SHEAT is set to zero in code in MSBSETVARS. 

When there is snow on the ground, albedo is the parameter ALBSN, otherwise it is the parameter ALB.

Estimation of net longwave radiation has been the subject of much research. BROOK90 uses [Brutsaert's (1982)](./literature.html) equation for effective clear sky emissivity (EFFEM).

The longwave equations and parameters are taken from [Brutsaert (1982)](./literature.html) and the net radiation extinction from [Shuttleworth and Wallace (1985)](./literature.html). 

The needed input is shown in the following table:

Input   |Description
--------|-----------------------------------
SLRAD   |solar radiation on slope (W/m2)
ALBEDO  |albedo
C1      |intercept of relation of solar radiation to sunshine duration
C2      |slope of relation of solar radiation to sunshine duration
C3      |longwave correction factor for overcast sky
TA      |air temperature (degC)
RATIO   |ratio of solar radiation on horizontal to potential insolation for day
EA      |vapor pressure, kPa
SHEAT   |average soil heat flux for the day (W/m2), usually 0
CR      |light extinction coefficient for projected LAI + SAI
LAI     |leaf area index (m2/m2)
SAI     |stem area index (m2/m2)

In the function will be calculated:

* SOLNET - net solar radiation (W/m2)
* EFFEM - effective emissivity from clear sky
* NOVERN - sunshine duration fraction of daylength
* CLDCOR - cloud cover correction to net longwave under clear sky
* LNGNET - net longwave radiation (W/m2)
* RN - net radiation (W/m2)

For EFFEM, the [Brutsaert (1982)](./literature.html) equation for effective clear sky emissivity will be used, because it's the most central. SIGMA is the Stefan-Boltzmann constant. 

With the help of this values the output of the function will be calculated:

* AA - available energy  (W/m2)
* ASUBS - available energy at ground (W/m2)

The available energy AA is the difference between net radiation and subsurface heat flux. The available energy at the ground is reduced exponentially by the CR, LAI and SAI, because of the radiation going through the canopy.

```{r}
AVAILEN<-function (SLRAD, ALBEDO, C1, C2, C3, TA, EA, RATIO, SHEAT, CR, LAI, SAI){
#local
 SOLNET <-0
 EFFEM<-0   
 NOVERN<-0  
 CLDCOR<-0  
 LNGNET<-0   
 RN  <-0    
#
SOLNET <- (1 - ALBEDO) * SLRAD
EFFEM <- 1.24 * (EA * 10 / (TA + 273.15)) ^ (1 / 7)
NOVERN <- (RATIO - C1) / C2
if (NOVERN > 1)  NOVERN <- 1
if (NOVERN < 0)  NOVERN <- 0
CLDCOR <- C3 + (1 - C3) * NOVERN
# emissivity of the surface taken as 1.0 to also account for reflected
LNGNET <- (EFFEM - 1) * CLDCOR * SIGMA * (TA + 273.15) ^ 4
RN <- SOLNET + LNGNET
AA <- RN - SHEAT
ASUBS <- RN * exp(-CR * (LAI + SAI)) - SHEAT
return(list(RN, AA, ASUBS))
}
```

### Function EQUIVSLP
In BROOK90 subroutine EQUIVSLP is called once to obtain parameters that depend on latitude, slope, and aspect of the surface being simulated.

The equivalent slope is defined as the location on the earth's surface where a horizontal surface is parallel to the given sloping surface. Following [Swift (1976)](./literature.html), L1 is the latitude of this "equivalent slope" (which is actually a horizontal surface), and L2 is the difference in hour angle (longitude) between the two locations. For any given slope and aspect, L1 and L2 need be found only once; they do not change over time. So they are calculated in EQUIVSLP at the beginning of B90, as:

* L1 = ASIN [COS(SLOPE) x SIN(LAT) + SIN(SLOPE) x COS(LAT) x COS(ASPECT)]
* L2 = ATAN {SIN(SLOPE) x SIN(ASPECT) / [COS(SLOPE) x COS(LAT) - SIN(SLOPE) x SIN(LAT) x COS(ASPECT)]}

with fixes for negative or zero denominator in L2, where SLOPE (ESLOPE), ASPECT, and latitude (LAT) are input parameters describing the location. All angles in the subroutine are in radians. 

The input of the function is included in the next table:

Input||Description
------|-------------------------
LAT   |latitude (radians (S neg))
SLOPE |slope (radians)
ASPECT|aspect (radians from N thru E)

The outputs of this function are:

* L1 - latitude of equivalent slope (radians)
* L2 - time shift of equivalent slope (radians)

To calculate these the external function ASIN is needed and instrinic are SIN, COS and ATN. For L2 the parameter D1 has to be calculated. It has to be set to 0.0000000001, if it has the value 0. Depending on the value of D1, L2 will be calculated in a different way. 

```{r}
EQUIVSLP<-function (LAT, SLOPE, ASPECT){
#Swift#s L1 and L2, Lee (3.31, 3.32)
#local
D1<-0
#
L1 <- ASINF(cos(SLOPE) * sin(LAT) + sin(SLOPE) * cos(LAT) * cos(ASPECT))
D1 <- cos(SLOPE) * cos(LAT) - sin(SLOPE) * sin(LAT) * cos(ASPECT)
if (D1 == 0)  D1 <- .0000000001
L2 <- atan(sin(SLOPE) * sin(ASPECT) / D1)
if (D1 < 0) L2 <- L2 + PI
return(list(L1, L2))
}

```

### Function SUNDS
Subroutine SUNDS uses functions FUNC3 and HAFDAY once a day to calculate daylength, potential radiation on a horizontal surface, and the ratio of potential radiation on the slope to that on a horizontal surface.

Several radiation-related variables depend only on day of the year and location. These are calculated in SUNDS, which is called once a day. 

SUNDS requires the solar constant (SCD), which is the radiation (W/m2) on a surface normal to the sun outside the atmosphere. It depends on day-of-the-year (DOY) to determine the earth-sun distance and is

* SCD = SC / (1 - .0167 x COS(.0172 x (DOY - 3))) ^ 2

where SC is the solar constant at the mean earth-sun distance ([Swift (1976)](./literature.html)). SC is set to 1367 W/m2 ([Lean (1991)](./literature.html)) and can not be changed.

The declination of the sun (DEC) is the angle by which the sun is above or below the plane of the earth's equator. DEC is zero at the equinoxes and +23.5° or -23.5 at the solstices. [Swift (1976)](./literature.html) gives the solar declination (radians) as:

* DEC = ASIN {0.39785 x SIN [4.86961 + 0.017203 x DOY + 0.033446 x SIN (6.224111 + 0.017202 x DOY)]}

The input to SUNDS includes:

Input|Description
-----|-----------------------------------------
LAT  |latitude (radians)
SLOPE|slope (radians)
DOY  |day of the year
L1   |latitude of equivalent slope (radians), from EQUIVSLP
L2   |time shift of equivalent slope (radians), from EQUIVSLP

For the calculation there are three constants necessary.

* WTOMJ (conversion factor: 0.0864 (MJ m-2 d-1)/(W/m2))
* PI
* SC (solar constant, 1367 W/m2)

Also external functions are needed and these are HAFDAY, FUNC3, RMIN, RMAX and ASIN. Instrinsic are COS and SIN. In this function many locals are defined and all receive the value 0:

Name  |Description
------|------------------------------------------------------------
I0SDAY|potential insolation on slope, map area basis (MJ/m2)
SCD   |solar constant for day (W/m2)
DEC   |declination of the sun (radians), zero at the equinoxes, +23.5° / -23.5 at the solstices
TWORIS|if two sunrises on slope
Temp  |temporary variable
T0    |hour angle of sunrise on horizontal (radians)
T1    |hour angle of sunset on horizontal
T2    |hour angle of sunrise on slope
T3    |hour angle of sunset on slope
T6    |hour angle of sunrise on equivalent slope
T7    |hour angle of sunset on equivalent slope
T8    |hour angle of second sunrise on slope
T9    |hour angle of second sunset on slope

T2 and T3 are used from the [Function HAFDAY](#function-hafday).
The calculated output of the function is:

* DAYLEN - Daylength (sun above horizontal) in fraction of day (d)
* I0HDAY - Potential insolation on horizontal surface (MJ/m2)
* SLFDAY - Ratio of potential insolation on slope to horizontal (map area)

The daylength will be evaluated and the function checks if there are two sunrises. Depending on if there are one or two sunrises on slope, the potential insolation on horizontal surface and the ratio of potential insolation on slope to horizontal have different results.

```{r}
SUNDS<-function (LAT, SLOPE, DOY, L1, L2, DAYLEN, I0HDAY, SLFDAY){
#local
I0SDAY <-0
SCD <-0  
DEC <-0 
TWORIS <-0 
Temp  <-0 
T0 <-0  
T1 <-0  
T2 <-0 
T3 <-0  
T6 <-0 
T7 <-0
T8 <-0  
T9 <-0   
#
SCD <- SC / (1 - .0167 * cos(.0172 * (DOY - 3))) ^ 2
DEC <- ASINF(.39785 * sin(4.868961 + .017203 * DOY + .033446 * sin(6.224111 + .017202 * DOY)))
Temp <- HAFDAY(LAT, DEC)
DAYLEN <- RMAXF(.0001, RMINF(.9999, Temp / PI))
# to avoid zero divides for 0 and 1
T1 <- Temp
T0 <- -Temp
Temp <- HAFDAY(L1, DEC)
T7 <- Temp - L2
T6 <- -Temp - L2
T3 <- RMINF(T1, T7)
T2 <- RMAXF(T0, T6)
  if (T3 < T2) {
    T2 <- 0
     T3 <- 0
  }
T6 <- T6 + 2 * PI
  if (T6 < T1) {
  T8 <- T6
  T9 <- T1
  TWORIS <- 1
  }
  T7 <- T7 - 2 * PI
  if (T7 > T0) {
    T8 <- T0
    T9 <- T7
    TWORIS <- 1
  }else{
  TWORIS <- 0
  }
if (TWORIS == 1) {   # two sunrises
  I0SDAY <- WTOMJ * SCD * (FUNC3(DEC, L2, L1, T3, T2) + FUNC3(DEC, L2, L1, T9, T8)) / cos(SLOPE)
# "daylength" on the slope = ((T3 - T2) + (T9 - T8)) / (2. * PI)
}else{    #  one sunrise
  I0SDAY <- WTOMJ * SCD * FUNC3(DEC, L2, L1, T3, T2) / cos(SLOPE)
# COS(SLOPE) adjusts from slope area to map area
# "daylength" on the slope = (T3 - T2) / (2. * PI)
}
  I0HDAY <- WTOMJ * SCD * FUNC3(DEC, 0, LAT, T1, T0)
  if (I0HDAY <= 0){
    SLFDAY <- 0
  }else{
    SLFDAY <- I0SDAY / I0HDAY
  }
  return(list(DAYLEN, I0HDAY, SLFDAY))
}
```

### Function FUNC3
The daily integral of potential insolation on a slope (I0SDAY, MJ/m2) is given by [Swift (1976)](./literature.html) as:

* I0SDAY = WTOMJ x SCD x FUNC3 (DEC, L2, L1, T3, T2)

where

* FUNC3 = (&pi;/ 2) {SIN(DEC) x SIN(L1) x (T3 - T2) + COS(DEC) x COS(L1) x [ SIN(T3 + L2) - SIN(T2 + L2)]} 

is a program function and T2 and T3 are the hour angles of sunrise and sunset on the slope, which are obtained from [Function HAFDAY](#function-hafday) using the latitude of the equivalent slope. FUNC3 has units of d-1 and WTOMJ is the conversion factor 0.0864 (MJ m-2 d-1) / (W m-2). The actual [Function SUNDS](#function-sunds) algorithm is more complicated than this because it must consider the possibility of two sunrises and sunsets on the slope in one day. The details of the algorithm are given by [Swift (1976)](./literature.html). Note that this algorithm assumes that the "opposing slope" is horizontal. In reality in mountainous terrain, the potential insolation is further reduced by any distant terrain that obscures the horizon more than the given slope itself does. The calculation of such obscuration is difficult and is outside the scope of BROOK90.

The daily integral of potential insolation on a horizontal surface (I0HDAY, MJ/m2) is found from the I0SDAY equation with L1 = LAT, L2 = 0, and T3 and T2 for a horizontal surface at LAT. The daylength (DAYLEN), which is the fraction of a day that the sun is above a horizontal horizon, is HAFDAY / &pi; where [Function HAFDAY](#function-hafday) is used with L = LAT. SLFDAY is the ratio of I0SDAY to I0HDAY. [SUNDS](#function-sunds) outputs DAYLEN, I0HDAY, and SLFDAY. 

FUNC3 is a function to calculate the daily integration for slope after [Swift (1976)](./literature.html), d. The necessary input data are:

Input |Description
------|----------------
DEC   |declination of the sun  (radians)
L2    |time shift of equivalent slope (radians)
L1    |latitude of equivalent slope (radians)
T3    |hour angle of sunset on slope
T2    |hour angle of sunrise on slope

For the calculation instrinsic are again SIN and COS. The output is FUnC3, which is the calculated daily integration for slope.

```{r}
FUNC3<-function(DEC, L2, L1, T3, T2){
#
FUnC3 <- (1 / (2 * 3.14159)) * (sin(DEC) * sin(L1) * (T3 - T2) + cos(DEC) * cos(L1) * (sin(T3 + L2) - sin(T2 + L2)))
return(FUnC3)
}
```

### Function HAFDAY
Daylength is calculated here as half-day length (HAFDAY), which is half the time interval between sunrise and sunset. The half-day length in radians is :

* HAFDAY = ACOS [-TAN(DEC) * TAN(L)]

where DEC is solar declination and L is either latitude or latitude of the equivalent slope. Function HAFDAY includes fixes for the poles or for the sun continuously above or below the horizon. 

In this case the external function ACOS is needed and ABS, SGN and TAN are instrinsic. Input arguments are:

Input |Description
------|----------------
LAT   |latitude  (radians (S neg))
DEC   |declination of the sun (radians)

The value ARG is calculated in the function to check if the sun stays below or above the horizon. If ARG is >= 1, the sun will stay below the horizon and if ARG is <= -1, the sun stays above the horizon. The output of the function is:

* HAFDAy - half daylength (radians)

```{r}
HAFDAY<-function (LAT, DEC){
#local
ARG<-0
#
if (abs(LAT) >= PI / 2)  LAT <- sign(LAT) * (PI / 2 - 0.01)
ARG <- -tan(DEC) * tan(LAT)
if (ARG >= 1) {
HAFDAy <- 0
}else if (ARG <= -1) {
HAFDAy <- PI
}else{
HAFDAy <- ACOSF(ARG)
}
return(HAFDAy)
}
```